Non-word repetition in Yélî Dnye

### 2020-09-17 manuscript.Rmd
add responses to AC comments

<!--M2A: I feel like there's a bit of a jump here from explaining what NWR is and what it's used for to saying we used it in the current study to test three things. We should probably have a paragraph justifying its use in this context, no?-->
<!-- a2m I added a sentence, is this what you were thinking of? -->

@@

<!-- a2m up to here it looks great. i am less sure about the next paragraph because it is unclear how it connects with our study... we didn't really take this into account when building the stimuli, right? -->

While all consonants can appear in the first syllable of the word, the second syllable onset consonant in bisyllabic roots is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43% and long: 31%) are more common than nasal ones (short: 4% and long: 22%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see @levinsoninprepydgrammar for details).

@@

Most speakers of the language grow up speaking it monolingually until they begin attending school around the age of 7; 
<!-- a2m no "at" here? *at* around the age? -->
school instruction is in English. While monolingual Yélî Dnye upbringing is common, multilingual families are not unusual, particularly in the region around the Catholic Mission---the same region in which the current data were collected---, where there is a higher incidence of married-in mothers from other islands [@brownIPchildrearing]. Children in these multilingual families grow up speaking Yélî Dnye and English, Tok Pisin, and/or some other language(s) from the region.

@@

While work on Yélî language development is growing [e.g., @brown2011cultural; @brown2014interactional; @brownIPchildrearing; @casillas2020early; @liszkowski2012prelinguistic], our knowledge of children's linguistic development is quite limited, and research into their phonological development in particular has only just begun [e.g., Bram's thesis; MC's CFH data]. What we do know about the early language environment is that, while child-directed speech is relatively infrequent throughout the day [@casillas2020early], shared caregiving practices and a near-universal fondness for social interaction with young children ensure that speech directed to children comes from all types of speakers: women, men, and other children [@bunceURcrosscultural; @casillas2020early]. While speech from women predominates in young children's speech environments on Rossel, as it does elsewhere [@bergelsoncasillas2019what; @bunceURcrosscultural], there is a significant and marked increase in child-directed speech from other children as infants get older [@bunceURcrosscultural; @casillas2020early]. This increase in child-directed speech from other children is attributed to the fact that, starting around age 2, children often spend large swaths of the day playing, swimming, and foraging for fruit, nuts, and shellfish in large (~10 people) independent and mixed-age child play groups [@brownIPchildrearing; @casillas2020early].
<!--M2A: I think a nice way to end this bit would be very briefly (~1-sentence) linking this input scenario to the relevance of looking at NWR (à la Cristia 2020 DevRev). What do you think?-->
<!-- a2m maybe we want to bring that up in the discussion? I haven't found a good place to introduce that paper, in part because right now we have these three purposes that seem sort of unconnected with the goals of that paper... -->

@@

The basic structure of NWR involves the participant hearing a non-word, like "bilik", which they need to repeat back without changing anything that is contrastive in the language. For instance, in English, a response of "bilig" or "pilik" will be scored as incorrect; a response "biilik", where the vowel is lengthened without change of quality would be scored as correct, because English does not have a pure length contrast. 
<!-- a2m added explanation of correct and incorrect for nwr task -->

@@

- Does the frequency of sounds across languages predict NWR? Are rarer sounds more often substituted by commoner sounds?
- How does score change as a function of item length in number of syllables?
- Is individual variation attributable to child age, sex, birth order, monolingual status, and/or parental education?
<!-- a2m we don't actually look at this: Are rarer sounds more often substituted by commoner sounds? -- shall we remove it? -->
<!--M2A: I'm imagining a reviewer who's knowlegeable about phonology and child development but who hasn't encountered NWR before. I think it would be beneficial to provide a concrete example of a nonword and an inexact repetition above so that they can conceive of what these questions imply (analytically) before we get into the methodological weeds.-->
<!-- a2m good point, but I think it goes in the previous section so I added it there -->

@@

myplot = ggplot(data = pl, aes(x = Age, y = Score, color = Language))+
  geom_point(aes(size = n, shape = Length))+
  geom_line(aes(linetype=Length)) + 
  theme(text = element_text(size = 16),
    legend.position="none", axis.text.x=element_text(colour=c("black", "black"))) +
  scale_color_manual(values = palette)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_x_continuous(name = "Age (years) ", limits = c(4,12), breaks = c(4,8,12))+

@@

# a2m I get a warning:
#   Warning message:
# Vectorized input to `element_text()` is not officially supported.
# Results may be unexpected or may change in future versions of ggplot2.

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items 'spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.
 <!--M2A: I didn't look for the SM yet, but probably best not to provide exact hamlet names-->
<!--a2m: good point! fixed-->

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether (see Participants section for exclusions). 
<!--M2A: do you know how many these were? oh! it's below. then maybe we can just put a (see Participants below), etc.-->
<!-- a2m done -->

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being "bilik", and imagine a response "bilig". This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20% (20% of phonemes were substituted or deleted).  
<!--M2A: A concrete scoring example would be a good idea here-->
<!-- a2m added! actually, I haven't changed anything but I've been hesitant about using levenshtein's distance because it's the complement of phoneme based scoring -- so shall we drop it?  -->

@@

```{r overall}
# M2A: This excludes a bunch of kids (anyone with NA on any of these participant features),
# whittling the summary down to 24 from 42 -- is that intended?
# Also all phon_score is 100 and all NLD is 0; does that stike you as fishy?
#a2m fixed and fixed!
aggregate(trnsc$correct*100,by=list(trnsc$id),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
overall=merge(overall, demo[,c("ID","Sex","Age","birthOrder","matEd")], by.x="Group.1",by.y="ID", all.x=T)
colnames(overall)<-c("ID","Word","Phoneme","NLD","Sex","Age","BrthO","matEd")

@@

<!-- a2m should we point out that the coefficient for cross-ling frequency is reduced, so indirectly corpus frequency does explain a little bit of variance? Or making things too complicated, and just say "so corpus frequency does not explain away these results" -->

@@

<!-- a2m I realized this section did not belong together with the xling freq section above - are you ok with this title? -->

@@

@@

<!-- a2m I added info above - is it weird to talk about this here? -->

@@

```{r bylength}
# M2A: this reduces the total number of kids represented from 42 to 40
#a2m fixed
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
by_length=merge(by_length, demo[,c("ID","Sex","Age","birthOrder","matEd")], by.x="Group.1",by.y="ID", all.x=T)
colnames(by_length)<-c("ID","Length","Word","Phoneme","NLD","Sex","Age","BrthO","matEd")

@@

Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(`r length_r$parameter`) = `r round(length_r$estimate,3)`.
<!-- a2m it doesn't make sense to do this, right, with only 3 items? -->

>>>>

<!--M2A: I feel like there's a bit of a jump here from explaining what NWR is and what it's used for to saying we used it in the current study to test three things. We should probably have a paragraph justifying its use in this context, no?-->
<!-- a2m I added a sentence, is this what you were thinking of? -->
<!-- M2A: Makes sense. I was thinking we could already hint at YD as interesting in particular bc of the phonology and not simply because it's a random one of many previously untested populations. I don't mind adding something like this in once I'm done with the demo check.-->

@@

<!-- a2m up to here it looks great. i am less sure about the next paragraph because it is unclear how it connects with our study... we didn't really take this into account when building the stimuli, right? -->
<!-- M2A: Happy to comment it out I agree it's not really relevant for our study in particular; just part of the phonological sketch. I was thinking about linguists when I wrote it up, but psychies probably are just ready to get on with the next part by now :) -->

<!--
While all consonants can appear in the first syllable of the word, the second syllable onset consonant in bisyllabic roots is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43% and long: 31%) are more common than nasal ones (short: 4% and long: 22%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see @levinsoninprepydgrammar for details).
-->

@@

Most speakers of the language grow up speaking it monolingually until they begin attending school around the age of 7; 
<!-- a2m no "at" here? *at* around the age? -->
<!-- M2A: my ear strongly prefers it without the "at", but it could go both ways. -->
school instruction is in English. While monolingual Yélî Dnye upbringing is common, multilingual families are not unusual, particularly in the region around the Catholic Mission---the same region in which the current data were collected---, where there is a higher incidence of married-in mothers from other islands [@brownIPchildrearing]. Children in these multilingual families grow up speaking Yélî Dnye and English, Tok Pisin, and/or some other language(s) from the region.

@@

While work on Yélî language development is growing [e.g., @brown2011cultural; @brown2014interactional; @brownIPchildrearing; @casillas2020early; @liszkowski2012prelinguistic], our knowledge of children's linguistic development is quite limited, and research into their phonological development in particular has only just begun [e.g., Bram's thesis; MC's CFH data]. What we do know about the early language environment is that, while child-directed speech is relatively infrequent throughout the day [@casillas2020early], shared caregiving practices and a near-universal fondness for social interaction with young children ensure that speech directed to children comes from all types of speakers: women, men, and other children [@bunceURcrosscultural; @casillas2020early]. While speech from women predominates in young children's speech environments on Rossel, as it does elsewhere [@bergelsoncasillas2019what; @bunceURcrosscultural], there is a significant and marked increase in child-directed speech from other children as infants get older [@bunceURcrosscultural; @casillas2020early]. This increase in child-directed speech from other children is attributed to the fact that, starting around age 2, children often spend large swaths of the day playing, swimming, and foraging for fruit, nuts, and shellfish in large (~10 people) independent and mixed-age child play groups [@brownIPchildrearing; @casillas2020early].
<!--M2A: I think a nice way to end this bit would be very briefly (~1-sentence) linking this input scenario to the relevance of looking at NWR (à la Cristia 2020 DevRev). What do you think?-->
<!-- a2m maybe we want to bring that up in the discussion? I haven't found a good place to introduce that paper, in part because right now we have these three purposes that seem sort of unconnected with the goals of that paper... -->
<!-- M2A: fine with me!-->

@@

The basic structure of NWR involves the participant hearing a non-word, like "bilik", which they need to repeat back without changing anything that is contrastive in the language. For instance, in English, a response of "bilig" or "pilik" will be scored as incorrect; a response "biilik", where the vowel is lengthened without change of quality would be scored as correct, because English does not have a pure length contrast. 
<!-- a2m added explanation of correct and incorrect for nwr task -->
<!--M2A: thanks!-->

@@

- Does the frequency of sounds across languages predict NWR? Are rarer sounds more often substituted by commoner sounds?
- How does score change as a function of item length in number of syllables?
- Is individual variation attributable to child age, sex, birth order, monolingual status, and/or parental education?
<!-- a2m we don't actually look at this: Are rarer sounds more often substituted by commoner sounds? -- shall we remove it? -->
<!--M2A: I'm imagining a reviewer who's knowlegeable about phonology and child development but who hasn't encountered NWR before. I think it would be beneficial to provide a concrete example of a nonword and an inexact repetition above so that they can conceive of what these questions imply (analytically) before we get into the methodological weeds.-->
<!-- a2m good point, but I think it goes in the previous section so I added it there -->
<!--M2A: thanks!-->

@@

myplot = ggplot(data = pl, aes(x = Age, y = Score, color = Language))+
  geom_point(aes(size = n, shape = Length))+
  geom_line(aes(linetype=Length)) + 
  theme(text = element_text(size = 16),
    legend.position="none", axis.text.x=element_text(colour="black")) +
  scale_color_manual(values = palette)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_x_continuous(name = "Age (years) ", limits = c(4,12), breaks = c(4,8,12))+

@@

# a2m I get a warning:
#   Warning message:
# Vectorized input to `element_text()` is not officially supported.
# Results may be unexpected or may change in future versions of ggplot2. 
# M2A: removed the vector since it wasn't needed; no more complaint!

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items 'spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.
 <!--M2A: I didn't look for the SM yet, but probably best not to provide exact hamlet names-->
<!--a2m: good point! fixed-->
<!-- M2A: thanks! -->

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether (see Participants section for exclusions). 
<!--M2A: do you know how many these were? oh! it's below. then maybe we can just put a (see Participants below), etc.-->
<!-- a2m done -->
<!-- M2A: thanks! -->

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being "bilik", and imagine a response "bilig". This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20% (20% of phonemes were substituted or deleted).  
<!--M2A: A concrete scoring example would be a good idea here-->
<!-- a2m added! actually, I haven't changed anything but I've been hesitant about using levenshtein's distance because it's the complement of phoneme based scoring -- so shall we drop it?  -->
<!-- M2A: three thoughts on this, which you can take as you like:
1) our audience might make this determination. segment-match score (phoneme-level accuracy) is probably more straightforward and might therefore be preferred by psychies
2) LD has the advantage of making D/I/S relevant as a next step in the analysis
3) we should definitely choose one, I agree, and report the other in supp mat tables
-->

@@

```{r overall}
# M2A: This excludes a bunch of kids (anyone with NA on any of these participant features),
# whittling the summary down to 24 from 42 -- is that intended?
# Also all phon_score is 100 and all NLD is 0; does that stike you as fishy?
#a2m fixed and fixed!
# M2A: thanks!
aggregate(trnsc$correct*100,by=list(trnsc$id),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
overall=merge(overall, demo[,c("ID","Sex","Age","birthOrder","matEd")], by.x="Group.1",by.y="ID", all.x=T)
colnames(overall)<-c("ID","Word","Phoneme","NLD","Sex","Age","BrthO","matEd")

@@

<!-- a2m should we point out that the coefficient for cross-ling frequency is reduced, so indirectly corpus frequency does explain a little bit of variance? Or making things too complicated, and just say "so corpus frequency does not explain away these results" -->
<!-- M2A: let's not go into it. because of the way glmer treats collinearity (https://psycnet.apa.org/record/2014-08495-005), it gets complicated quickly in even giving an accurate report of what's going on -->

@@

<!-- a2m I realized this section did not belong together with the xling freq section above - are you ok with this title? -->
<!-- M2A: The title is a little long and maybe less informative than it could be. How about something like, "Patterns in NWR mispronunciations"-->

@@

<!--M2A: coming back to these three comments below after I manage another full read-through (after I check demo) -->

@@

<!-- a2m I added info above - is it weird to talk about this here? -->
<!-- M2A: I don't think it's a weird place to add the idea, though I might shorten it a bit when I go through on a writing pass :) -->

@@

```{r bylength}
# M2A: this reduces the total number of kids represented from 42 to 40
#a2m fixed
# M2A: thanks!
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
by_length=merge(by_length, demo[,c("ID","Sex","Age","birthOrder","matEd")], by.x="Group.1",by.y="ID", all.x=T)
colnames(by_length)<-c("ID","Length","Word","Phoneme","NLD","Sex","Age","BrthO","matEd")

@@

Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(`r length_r$parameter`) = `r round(length_r$estimate,3)`.
<!-- a2m it doesn't make sense to do this, right, with only 3 items? -->
<!-- M2A: you mean with only three lengths? I guess I would prefer a little regression -->
